Interpolation of Sparse High-Dimensional Data\thanks {This work was supported by the National Science Foundation Grant CNS-1565314.}

### 2019-03-07 Paper_[2019-02]_Numerical_Algorithms/NA2019b.tex
Before first round of revisions.e

In this work, multivariate interpolation is defined given a closed convex subset $Y$ of a metrizable topological vector space with metric $s$, some function $f:\mathbb{R}^d \rightarrow Y$ and a set of points $X = \bigl\{x^{(1)}$,$\ldots$,$x^{(n)}\bigr\} \subset \mathbb{R}^d,$ along with associated function values $f\bigl(x^{(i)}\bigr)$. The goal is to construct an approximation $\hat f: \mathbb{R}^d \rightarrow Y$ such that $\hat f\bigl(x^{(i)}\bigr) = f\bigl(x^{(i)}\bigr)$ for all $i = 1$,$\ldots$,$n$. It is often the case that the form of the true underlying function $f$ is unknown, however it is still desirable to construct an approximation $\hat f$ with small approximation error at $y \notin X$. The two metric spaces that will be discussed in this work are the real numbers with metric $s(x,y) = |x-y|$, and the set of cumulative distribution functions (CDFs) with the Kolmogorov-Smirnov (KS) statistic as metric.

@@

The remainder of this paper is organized as follows. Sections \ref{sec:regression} and \ref{sec:interpolation} present the multivariate models. Section \ref{sec:error} presents the error measuring methodology that is used for collecting empirical results. Section \ref{sec:theory} presents the theoretical error bounds for interpolation. Section \ref{sec:data} presents the data sets for empirical approximation analysis and presents the approximation results for all models on these data sets. Section \ref{sec:discussion} analyzes and discusses the results, their implications, and their limitations. Finally, Section \ref{sec:conclusion} concludes.

@@

\subsection{Support Vector Regressor}
\label{sec:svr}
Support vector machines are a common method used in machine learning classification tasks that can be adapted for the purpose of regression \cite{basak2007support}. How to build a support vector regressor (SVR) is beyond the scope of this summary, but the resulting functional fit $p : \mathbb{R}^d \rightarrow \mathbb{R}$ has the form
$$ p(x)  = \sum_{i=1}^{n}a_i K\bigl(x,x^{(i)}\bigr) + b ,$$
where $K$ is the selected kernel function, $a_i \in \mathbb{R}^n$, $b \in \mathbb{R}$ are coefficients to be solved for simultaneously. The computational complexity of the SVR is $\mathcal{O}(n^2dm)$, with $m$ being determined by the minimization convergence criterion. This paper uses the scikit-learn SVR \cite{scikit-learn} with a polynomial kernel function.

\subsection{Multilayer Perceptron Regressor}
\label{sec:mlp}
The neural network is a well studied and widely used method for both regression and classification tasks \cite{rumelhart1988learning,hornik1989multilayer}. When using the rectified linear unit (ReLU) activation function \cite{dahl2013improving} and fitting the model with stochastic gradient descent (SGD) or BFGS minimization techniques \cite{goh2017why,moller1993scaled,robbins1951stochastic}, the model built by a multilayer perceptron uses layers $l : \mathbb{R}^{i} \rightarrow \mathbb{R}^{j}$ defined by
$$ l(u) = \bigl( u^t W_l \bigr)_+ ,$$
where $W_l$ is the $i$ by $j$ weight matrix for layer $l$. In this form, the multilayer perceptron (MLP) produces a piecewise linear model of the input data. The computational complexity of training a multilayer perceptron is $\mathcal{O}(n d m)$, where $m$ is determined by the sizes of the layers of the network and the stopping criterion of the minimization used for finding weights. This paper uses a MLP built from Keras and Tensorflow to perform regression \cite{chollet2015keras,tensorflow2015-whitepaper} with ten hidden layers each having thirty two nodes (a total of approximately ten thousand parameters), ReLU activation, and five thousand epochs of SGD for training.

@@

The modified Shepard method used here (ShepMod) generates a continuous approximation based on Euclidean distance and resembles a nearest neighbor interpolant \cite{cover1967nearest}. This model is a type of \textit{metric interpolation}, also called a Shepard method \cite{shepard1968two,gordon1978shepard}. The interpolant has the form
$$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)f\bigl(x^{(k)}\bigr)}{\sum\limits_{k=1}^{n}W_k(x)} ,$$
where $W_k(x) \propto \bigl\|x - x^{(k)}\bigr\|_2^{-2}$ and decreases to zero at a radius $r_k \in \mathbb{R}$ such that exactly $d+1$ other points $x^{(j)}$, $j \not = k$ are inside the closed Euclidean ball of radius $r_k$ about $x^{(k)}$. The computational complexity of ShepMod is $\mathcal{O}(n^2d)$. This paper uses a Fortran 95 implementation of ShepMod derived from SHEPPACK \cite{thacker2010algorithm}.

\subsection{Linear Shepard}
The linear Shepard method (LSHEP) is a blending function using local linear interpolants, a special case of the general Shepard algorithm \cite{thacker2010algorithm}. The interpolant has the form
$$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)P_k(x)}{\sum\limits_{k=1}^{n}W_k(x)} ,$$
where $W_k(x)$ is a locally supported weighting function and $P_k(x)$ is a local linear approximation to the data satisfying $P_k\bigl(x^{(k)}\bigr) = f\bigl(x^{(x)}\bigr)$. The computational complexity of LSHEP is $\mathcal{O}(n^2d^3)$. This paper uses the Fortran 95 implementation of LSHEP in SHEPPACK \cite{thacker2010algorithm}.

@@

A well-studied technique for classification and approximation is the nearest neighbor algorithm \cite{cover1967nearest}. Nearest neighbor inherently utilizes the convex region $C^{x^{(i)}}$ (Voronoi cell \cite{dirichlet1850reduction}) consisting of all points closer to $x^{(i)}$ than to any other point $x^{(j)}$. The Voronoi model smooths the nearest neighbor approximation by utilizing the Voronoi cells to define support via a generic basis function $v: \mathbb{R}^d \rightarrow \mathbb{R}_+$ given by
$$ v^{x^{(i)}}(y) = \left(1 - \frac{\bigl\|y - x^{(i)}\bigr\|_2}{2 \ h\bigl(y - x^{(i)} \mid x^{(i)}\bigr)} \right)_+, $$
where $h\bigl(w \mid x^{(i)}\bigr)$ is the distance between $x^{(i)}$ and the boundary of the Voronoi cell $C^{x^{(i)}}$ in the direction $w$. $v^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $v^{x^{(i)}}$ has local support, giving the interpolated value
$$ f(y) = \frac{\sum\limits_{i=1}^n v^{x^{(i)}}(y) f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n v^{x^{(i)}}(y)}, $$
where $0 \leq v^{x^{(i)}}(y) \leq 1$. The computational complexity of interpolation via this Voronoi model is $\mathcal{O}(n^2 d)$. All of the approximations are an interpolant involving a convex combination of known function values $f\bigl(x^{(i)}\bigr)$.

@@

A hurdle when modeling function-valued outputs such as cumulative distribution functions (CDFs) or probability density functions (PDFs) is that certain properties must be maintained. It is necessary that a PDF $f: \mathbb{R} \rightarrow \mathbb{R}$ have the properties $f(x) \geq 0$ and $\int_{-\infty}^{\infty}f(x)dx = 1$. However, for a CDF $F: \mathbb{R} \rightarrow \mathbb{R}$ the properties are $F(x) \in [0,1]$ and $F(x)$ is absolutely continuous and nondecreasing. This work utilizes the fact that a convex combination of CDFs (or PDFs) results in a valid CDF. Given $G(x) = \sum_{i}w_i F_i(x)$, $\sum_{i} w_i = 1$, $w_i \geq 0$, and each $F_i$ is a valid CDF, $G$ must also be a valid CDF. A demonstration of how this is applied can be seen in Figure \ref{fig:prediction-example}.

@@

This section presents the theoretical results bounding the error of Delaunay interpolation. The error analysis relies on the Delaunay interpolation for two reasons: (1) second order results can be obtained utilizing a Lipschitz constant on the gradient of a function, rather than standard Lipschitz bounds and (2) multiple other interpolants analyzed compute predictions as convex combinations of observed function values, which may allow for straight forward extensions of this error bound.

@@

In summary, the approximation error of a Delaunay interpolant tends quadratically towards zero when approaching observed data only when the diameter of the interpolating simplex is also reduced at a proportional rate. Without the incorporation of additional observations, only linear convergence to the true function is be achieved. Notice that the approximation error is largely determined by the spacing of observed data. Predictions made by simplices whose vertices are not well-spaced (i.e. have large diameter, or are nearly contained in a hyperplane) have higher error. In light of this error bound, an empirical evaluation of presented algorithms follows.

@@

The theoretical results constructed in Section \ref{sec:theory} for Delaunay interpolation are promising, however they are difficult to interpret in context with the other approximation algorithms that do not have similar known uniform error bounds. For that reason, this paper utilizes five different data sets of varying dimension and application to construct approximations and compare the accuracy of different techniques.

In the following five subsections the sources and targets of each test data set are described, as well as challenges and limitations related to approximating these data. The distribution of response values being modeled is presented followed by the distribution of approximation errors for each algorithm.

@@

All approximations are constructed using $10$-fold cross validation as described in \cite{kohavi1995study}. This approach requires the division of data randomly into $10$ partitions (the same partitions are used by each algorithm). An algorithm is then evaluated by constructing an approximation over each set of $9$ partitions, predicting the values of the final partition. As a result, each observed data point is used in the construction of $9$ different approximations and is approximated exactly once. The $k$-fold cross validation method is data-efficient and provides an unbiased estimate of the expected prediction error \cite{kohavi1995study}, however it should be noted that neither this method nor others can provide a universally unbiased estimator for the variance of prediction error \cite{bengio2004no}.

In addition to the figures displaying approximation results for each data set, tables of accompanying numerical results are located in the Appendix. All of the test data sets capture underlying functions that are almost certainly stochastic. As described in Section \ref{sec:introduction}, regression techniques appear most appropriate for these problems. However, in high dimension data grows exponentially more sparse. Given sparse data, regressors tend towards interpolation.

%% -------------------------------------------------------------------
\subsection{Forest Fire ($n = 504, d = 12$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest First Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of forest fire area burned under recorded weather conditions. The data is presented on a natural log scale because most values are small with exponentially fewer fires on record that burn large areas.}
  \label{fig:hist-forest-fire}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area that would be burned given environment conditions. The data is broken up into $10$ partitions and each set of $9$ partitions is used to construct a model that approximates the remaining partition. This results in exactly one prediction from each algorithm for each data point. These boxes depict the median (middle bar), median $95\%$ confidence interval (cones), quartiles (box edges), fences at $3/2$ interquartile range (whiskers), and outliers (dots) of absolute prediction error for each model. Similar to figure \ref{fig:hist-forest-fire}, the errors are presented on a natural log scale. The complementing numerical data to this figure is provided in table \ref{table:error-forest-fire} in the Appendix.}
  \label{fig:error-forest-fire}
\end{figure}

The forest fire data set \cite{cortez2007data} describes the area of Montesinho park burned over months of the year along with environmental conditions. The twelve dimensions being used to model burn area are the $x$ and $y$ spatial coordinates of burns in the park, month of year (mapped to a 2D circle), the FFMC, DMC, DC, and ISI indices (see source for details), the temperature, relative humidity, wind speed, and outdoor rain. The original analysis of this data set demonstrated it to be difficult to model, likely due to the skew in response values.

As suggested by figure \ref{fig:error-forest-fire}, the SVR has the lowest absolute prediction errors for 80\% of the data, with ShepMod and BoxSpline being the nearest overall competitors. The effectiveness of SVR on this data suggests the underlying function can be defined by relatively few parameters, as well as the importance of capturing the low-burn-area data points.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Parkinson's Telemonitoring ($n = 5875, d = 19$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of the Parkinson's patient total UPDRS clinical scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score given audio features from patients' home life. The data is broken up into $10$ partitions and each set of $9$ partitions is used to construct a model that approximates the remaining partition. This results in exactly one prediction from each algorithm for each data point. These boxes depict the median (middle bar), median $95\%$ confidence interval (cones), quartiles (box edges), fences at $3/2$ interquartile range (whiskers), and outliers (dots) of absolute prediction error for each model. The complementing numerical data to this is provided in table \ref{table:error-parkinsons} in the Appendix.}
  \label{fig:error-parkinsons}
\end{figure}

The second data set for evaluation \cite{tsanas2010accurate} is derived from a speech monitoring study with the intent to automatically estimate Parkinson's disease symptom development in Parkinson's patients. The function to be approximated is a time-consuming clinical evaluation measure referred to as the UPDRS score. The total UPDRS score given by a clinical evaluation is estimated through 16 real numbers generated from biomedical voice measures of in-home sound recordings.

Figure \ref{fig:error-parkinsons} shows the ShepMod algorithm has the lowest minimum, first quartile, and median of absolute errors for this problem, while providing the best prediction 66\% of the time. The MLP has the lowest third quartile and provides the best prediction for 32\% of approximations. The dominance of ShepMod may be in part to regular-interval total UPDRS scores provided by clinicians, favoring a nearest-neighbor strategy of prediction.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Australian Daily Rainfall Volume ($n = 2609, d = 23$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Daily Rainfall in Sydney};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Above is a histogram of daily rainfall in Sydney, Australia. The data is presented on a natural log scale because the frequency of larger amounts of rainfall is significantly less. There is a peak in occurrence of the value $0$, which has a notable effect on the resulting model performance.}
  \label{fig:hist-weather}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Sydney Rainfall Tomorrow Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of rainfall expected on the next calendar day given various sources of local meteorological data. The data is broken up into $10$ partitions and each set of $9$ partitions is used to construct models that approximates the remaining partition. This results in exactly one prediction from each algorithm for each data point. These boxes depict the median (middle bar), median $95\%$ confidence interval (cones), quartiles (box edges), fences at $3/2$ interquartile range (whiskers), and outliers (dots) of absolute prediction error for each model. The errors are presented on a natural log scale, mimicking the presentation in figure \ref{fig:hist-weather}. The complementing numerical data to this figure is provided in table \ref{table:error-weather} in the Appendix.}
  \label{fig:error-weather}
\end{figure}

The third data regarding the total daily rainfall in Sydney, Australia \cite{williams2009rattle} provides a slightly higher dimension challenge for the interpolants and regressors. This data is composed of many meteorological readings from the region on a day including: min and max temperatures, sunshine, wind speed directions (converted to coordinates on circle), wind speeds, and humidities throughout the day, day of the year (converted to coordinates on circle), and the model must predict the amount of rainfall tomorrow.

While figure \ref{fig:hist-weather} makes MARS look far better than other techniques, it only provides the best prediction for 11\% of points. The MLP has the lowest absolute error for 56\% of points and LSHEP is best for 28\%. MARS likely achieves such a low first quartile due to the high occurrence of the value 0 in the data.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Credit Card Transaction Amount ($n = 5562, d = 28$)}

>>>>

In this work, multivariate interpolation is defined given a closed convex
subset $Y$ of a metrizable topological vector space with metric $s$, some
function $f:\mathbb{R}^d \rightarrow Y$ and a set of points $X =
\bigl\{x^{(1)}$, $\ldots$, $x^{(n)}\bigr\} \subset \mathbb{R}^d,$ along with
associated function values $f\bigl(x^{(i)}\bigr)$. The goal is to construct
an approximation $\hat f: \mathbb{R}^d \rightarrow Y$ such that $\hat
f\bigl(x^{(i)}\bigr) = f\bigl(x^{(i)}\bigr)$ for all $i = 1$, $\ldots$, $n$.
It is often the case that the form of the true underlying function $f$ is
unknown, however it is still desirable to construct an approximation $\hat
f$ with small approximation error at $y \notin X$. The two metric spaces
that will be discussed in this work are the real numbers with metric $s(x,y)
= |x-y|$, and the set of cumulative distribution functions (CDFs) with the
Kolmogorov-Smirnov (KS) statistic as metric.

@@

The remainder of this paper is organized as follows. Sections
\ref{sec:regression} and \ref{sec:interpolation} present the multivariate
models. Section \ref{sec:error} gives the error measuring methodology
that is used for collecting empirical results. Section \ref{sec:theory}
contains the theoretical error bounds for interpolation. Section
\ref{sec:data} presents the data sets for empirical approximation analysis
and the approximation results for all models on these data sets.
Section \ref{sec:discussion} analyzes and discusses the results, their
implications, and their limitations. Finally, Section \ref{sec:conclusion}
concludes.

@@

\subsection{Support Vector Regressor}
\label{sec:svr}
Support vector machines are a common method used in machine learning
classification tasks that can be adapted for the purpose of regression
\cite{basak2007support}. How to build a support vector regressor (SVR) is
beyond the scope of this summary, but the resulting functional fit $p :
\mathbb{R}^d \rightarrow \mathbb{R}$ has the form
 $$ p(x)  = \sum_{i=1}^{n}a_i K\bigl(x,x^{(i)}\bigr) + b ,$$
where $K$ is the selected kernel function, $a \in \mathbb{R}^n$, $b \in
\mathbb{R}$ are coefficients to be solved for simultaneously. The
computational complexity of the SVR is $\mathcal{O}(n^2dm)$, with $m$ being
determined by the minimization convergence criterion. This paper uses the
scikit-learn SVR \cite{scikit-learn} with a polynomial kernel function.

\subsection{Multilayer Perceptron Regressor}
\label{sec:mlp}
The neural network is a well studied and widely used method for both
regression and classification tasks
\cite{rumelhart1988learning,hornik1989multilayer}. When using the rectified
linear unit (ReLU) activation function \cite{dahl2013improving} and fitting
the model with stochastic gradient descent (SGD) or BFGS minimization
techniques \cite{goh2017why,moller1993scaled,robbins1951stochastic}, the
model built by a multilayer perceptron uses layers $l : \mathbb{R}^{i}
\rightarrow \mathbb{R}^{j}$ defined by
 $$ l(u) = \bigl( u^t W_l \bigr)_+ ,$$
where $W_l$ is the $i\times j$ weight matrix for layer $l$. In this form,
the multilayer perceptron (MLP) produces a piecewise linear model of the
input data. The computational complexity of training a multilayer perceptron
is $\mathcal{O}(n d m)$, where $m$ is determined by the sizes of the layers
of the network and the stopping criterion of the minimization used for
finding weights. This paper uses a MLP built from Keras and Tensorflow to
perform regression \cite{chollet2015keras,tensorflow2015-whitepaper} with
ten hidden layers each having thirty two nodes (a total of approximately
ten thousand parameters), ReLU activation, and five thousand epochs of SGD
for training.

@@

The modified Shepard method used here (ShepMod) generates a continuous
approximation based on Euclidean distance and resembles a nearest neighbor
interpolant \cite{cover1967nearest}. This model is a type of \textit{metric
interpolation}, also called a Shepard method
\cite{shepard1968two,gordon1978shepard}. The interpolant has the form
 $$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)f\bigl(x^{(k)}\bigr)}
 {\sum\limits_{k=1}^{n}W_k(x)} ,$$
where $W_k(x) = \bigl(r_k - \bigl\|x - x^{(k)}\bigr\|\bigr)_+ \big/
\bigl\|x - x^{(k)}\bigr\|_2^2$ and $r_k \in \mathbb{R}$ is the smallest
radius such that at least $d+1$ other points $x^{(j)}$, $j \not = k$, are
inside the closed Euclidean ball of radius $r_k$ about $x^{(k)}$. The
computational complexity of ShepMod is $\mathcal{O}(n^2d)$. This paper
uses a Fortran 95 implementation of ShepMod derived from SHEPPACK
\cite{thacker2010algorithm}.

\subsection{Linear Shepard}
The linear Shepard method (LSHEP) is a blending function using local linear
interpolants, a special case of the general Shepard algorithm
\cite{thacker2010algorithm}. The interpolant has the form
 $$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)P_k(x)}
 {\sum\limits_{k=1}^{n}W_k(x)} ,$$
where $W_k(x)$ is a locally supported weighting function and $P_k(x)$ is
a local linear approximation to the data satisfying $P_k\bigl(x^{(k)}\bigr)
= f\bigl(x^{(x)}\bigr)$. The computational complexity of LSHEP is
$\mathcal{O}(n^2d^3)$. This paper uses the Fortran 95 implementation of
LSHEP in SHEPPACK \cite{thacker2010algorithm}.

@@

A well-studied technique for classification and approximation is the nearest
neighbor algorithm \cite{cover1967nearest}. Nearest neighbor inherently
utilizes the convex region $C^{x^{(i)}}$ (Voronoi cell
\cite{dirichlet1850reduction}) consisting of all points closer to $x^{(i)}$
than to any other point $x^{(j)}$. The Voronoi model smooths the nearest
neighbor approximation by utilizing the Voronoi cells to define support
via a generic basis function $v: \mathbb{R}^d \rightarrow \mathbb{R}_+$
given by
 $$ v^{x^{(i)}}(y) = \left(1 - \frac{\bigl\|y - x^{(i)}\bigr\|_2} {2 \
 h\bigl(y - x^{(i)} \mid x^{(i)}\bigr)} \right)_+, $$
where $h\bigl(w \mid x^{(i)}\bigr)$ is the distance between $x^{(i)}$ and
the boundary of the Voronoi cell $C^{x^{(i)}}$ in the direction $w$.
$v^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $v^{x^{(i)}}$ has local
support, giving the interpolated value
 $$ f(y) = \frac{\sum\limits_{i=1}^n v^{x^{(i)}}(y)
 f\bigl(x^{(i)}\bigr)}{\sum\limits_{i=1}^n v^{x^{(i)}}(y)}, $$
where $0 \leq v^{x^{(i)}}(y) \leq 1$. The computational complexity of
interpolation via this Voronoi model is $\mathcal{O}(n^2 d)$. All of the
approximations are an interpolant involving a convex combination of known
function values $f\bigl(x^{(i)}\bigr)$.

@@

A hurdle when modeling function-valued outputs such as cumulative
distribution functions (CDFs) or probability density functions (PDFs) is
that certain properties must be maintained. It is necessary that a PDF $f:
\mathbb{R} \rightarrow \mathbb{R}$ have the properties $f(x) \geq 0$ and
$\int_{-\infty}^{\infty}f(x)dx = 1$. However, for a CDF $F: \mathbb{R}
\rightarrow \mathbb{R}$ the properties are $F(x) \in [0,1]$ and $F(x)$ is
absolutely continuous and nondecreasing. This work utilizes the fact that
a convex combination of CDFs (or PDFs) results in a valid CDF (or PDF).
Given $G(x) = \sum_{i}w_i F_i(x)$, $\sum_{i} w_i = 1$, $w_i \geq 0$, and
each $F_i$ is a valid CDF, $G$ must also be a valid CDF. A demonstration
of how this is applied can be seen in Figure \ref{fig:prediction-example}.

@@

This section presents the theoretical results bounding the error of Delaunay
interpolation. The error analysis relies on the Delaunay interpolation for
two reasons: (1) second order results can be obtained utilizing a Lipschitz
constant on the gradient of a function, rather than standard Lipschitz
bounds, and (2) multiple other interpolants analyzed compute predictions
as convex combinations of observed function values, which may allow for
straightforward extensions of this error bound.

@@

In summary, the approximation error of a Delaunay interpolant tends
quadratically towards zero when approaching observed data only when the
diameter of the interpolating simplex is also reduced at a proportional
rate. Without the incorporation of additional observations, only linear
convergence to the true function is be achieved. Notice that the
approximation error is largely determined by the spacing of observed data.
Predictions made by simplices whose vertices are not well-spaced (i.e.,
have large diameter, or are nearly contained in a hyperplane) have higher
error. In light of this error bound, an empirical evaluation of the
presented algorithms follows.

@@

The theoretical results constructed in Section \ref{sec:theory} for Delaunay
interpolation are promising, however they are difficult to interpret in
context with the other approximation algorithms that do not have similar
known uniform error bounds. For that reason, this paper utilizes five
different application data sets of varying dimension to construct
approximations and compare the accuracy of different techniques.

In the following five subsections the sources and targets of each test
data set are described, as well as challenges and limitations related to
approximating these data. The distribution of response values being modeled
is presented followed by the distribution of approximation errors for each
algorithm. The plots for all five data sets have the same statistical format.

@@

All approximations are constructed using 10-fold cross validation as
described in \cite{kohavi1995study}. This approach requires the partition
of data randomly into ten subsets (the same subsets are used by
each algorithm). An algorithm is then evaluated by constructing an
approximation over each set of nine subsets, predicting the values of
the omitted subset. As a result, each observed data point is used in the
construction of nine different approximations and is approximated exactly
once. The $k$-fold cross validation method is data-efficient and provides
an unbiased estimate of the expected prediction error \cite{kohavi1995study},
however it should be noted that neither this method nor others can provide
a universally unbiased estimator for the variance of prediction error
\cite{bengio2004no}.

In addition to the figures displaying approximation results for each data
set, tables of accompanying numerical results are located in the Appendix.
All of the test data sets capture underlying functions that are almost
certainly stochastic. As described in Section \ref{sec:introduction},
regression techniques appear most appropriate for these problems. However,
typically data grows exponentially more sparse with increasing
dimension. Given sparse data, regressors tend towards interpolation.

%% -------------------------------------------------------------------
\subsection{Forest Fire ($n = 504$, $d = 12$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Forest Fire Area Burned};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of forest fire area burned under recorded weather
conditions. The data is presented on a $\log_{10}$ scale because most values
are small with exponentially fewer fires on record that burn large areas.}
  \label{fig:hist-forest-fire}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-forest-fire.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Area Burned Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of area that
  would be burned given environmental conditions. The data is partitioned
  into ten subsets and each set of nine subsets is used to construct a
  model that approximates the omitted subset. This results in exactly one
  prediction from each algorithm for each data point. These boxes depict
  the median (middle bar), median $95\%$ confidence interval (cones),
  quartiles (box edges), fences at $3/2$ interquartile range (whiskers),
  and outliers (dots) of absolute prediction error for each model. Similar
  to Figure \ref{fig:hist-forest-fire}, the errors are presented on a
  $\log_{10}$ scale. The numerical data corresponding to this figure is
  provided in Table \ref{table:error-forest-fire} in the Appendix.}
\label{fig:error-forest-fire} \end{figure}

The forest fire data set \cite{cortez2007data} describes the area of
Montesinho park burned over months of the year along with environmental
conditions. The twelve dimensions being used to model burn area are the
$x$ and $y$ spatial coordinates of burns in the park, month of year (mapped
to a 2D circle), the FFMC, DMC, DC, and ISI indices (see source for details),
the temperature, relative humidity, wind speed, and outdoor rain. The
original analysis of this data set demonstrated it to be difficult to
model, likely due to the skew in response values.  ***Missing 12th
variable above*****

As suggested by Figure \ref{fig:error-forest-fire}, the SVR has the lowest
absolute prediction errors for 80\% of the data, with MLP and Delaunay
being the nearest overall competitors. The effectiveness of SVR on this
data suggests the underlying function can be defined by relatively few
parameters, as well as the importance of capturing the low-burn-area data
points. 
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Parkinson's Telemonitoring ($n = 5875$, $d = 19$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Parkinson's Total UPDRS Score};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of the Parkinson's patient total UPDRS clinical
scores that will be approximated by each algorithm.}
  \label{fig:hist-parkinsons}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-parkinsons.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Total UPDRS Score Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the total UPDRS score
  given audio features from patients' home life, using 10-fold cross
  validation.  These boxes depict the median (middle bar), median $95\%$
  confidence interval (cones), quartiles (box edges), fences at $3/2$
  interquartile range (whiskers), and outliers (dots) of absolute prediction
  error for each model. The numerical data corresponding to this figure
  is provided in Table \ref{table:error-parkinsons} in the Appendix.}
 \label{fig:error-parkinsons}
\end{figure}

The second data set for evaluation \cite{tsanas2010accurate} is derived
from a speech monitoring study with the intent to automatically estimate
Parkinson's disease symptom development in Parkinson's patients. The
function to be approximated is a time-consuming clinical evaluation measure
referred to as the UPDRS score. The total UPDRS score given by a clinical
evaluation is estimated through 19 real numbers generated from biomedical
voice measures of in-home sound recordings.

Figure \ref{fig:error-parkinsons} shows the ShepMod algorithm has the
lowest minimum, first quartile, and median of absolute errors for this
problem, while providing the best prediction 66\% of the time. The MLP has
the lowest third quartile and provides the best prediction for 32\% of
approximations. The dominance of ShepMod may be due in part to
regular-interval total UPDRS scores provided by clinicians, favoring a
nearest-neighbor strategy of prediction.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Australian Daily Rainfall Volume ($n = 2609$, $d = 23$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Daily Rainfall in Sydney};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of daily rainfall in Sydney, Australia, presented on a
$\log_{10}$ scale because the frequency of larger amounts of rainfall
is significantly less. There is a peak in occurrence of the value $0$,
which has a notable effect on the resulting model performance.}
  \label{fig:hist-weather}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.8\textwidth,]{Figures/error-box-weather.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Sydney Rainfall Tomorrow Error};
  \end{tikzpicture}
  \caption{All models are applied to approximate the amount of rainfall
  expected on the next calendar day given various sources of local
  meteorological data, using 10-fold cross validation.
  These boxes depict the median (middle bar), median $95\%$ confidence
  interval (cones), quartiles (box edges), fences at $3/2$ interquartile
  range (whiskers), and outliers (dots) of absolute prediction error for
  each model. The errors are presented on a $\log_{10}$ scale, mimicking
  the presentation in Figure \ref{fig:hist-weather}. The 
  numerical data corresponding to this figure is provided in Table
  \ref{table:error-weather} in the Appendix.}
\label{fig:error-weather}
\end{figure}

The third data set for the total daily rainfall in Sydney, Australia
\cite{williams2009rattle} provides a slightly higher dimensional challenge
for the interpolants and regressors. This data is composed of many
meteorological readings from the region in a day including: min and max
temperatures, sunshine, wind speed directions (converted to coordinates
on a circle), wind speeds, and humidities throughout the day, day of the
year (converted to coordinates on a circle), and the model must predict the
amount of rainfall tomorrow.

While Figure \ref{fig:hist-weather} makes MARS look far better than other
techniques, it only provides the best prediction for 11\% of points. The
MLP has the lowest absolute error for 56\% of points and LSHEP is best for
28\%.  MARS likely achieves such a low first quartile due to the high
occurrence of the value zero in the data.
%% -------------------------------------------------------------------

%% -------------------------------------------------------------------
\subsection{Credit Card Transaction Amount ($n = 5562$, $d = 28$)}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.75\textwidth,]{Figures/raw-histogram-credit-card.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {Credit Card Transaction Amount};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{Histogram of credit card transaction amounts, presented on a
  $\log_{10}$ scale. The data contains a notable frequency peak around
  $\$1$ transactions. Fewer large purchases are made, but some large
  purchases are in excess of five orders of magnitude greater than the
  smallest purchases.}
